Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-11 paper.tex
Describe rule group mechanism.

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?"
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` },
    ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
Narrative is the core of the Nell interface.  It is important that
Nell's story be modular and extensible.  We use a rule-based story
model with programmable conflict resolution to provide this ability.
Figure~\ref{fig:rules} shows an excerpt of the story module driving
the interaction described in the introduction to this paper.

Stories consist of a sequence of \textit{scenes}, in addition to
characters, locations, and activity information.  Scenes are ``goals''
in the STRIPS formalism~\cite{strips}, and consist of
action descriptions, preconditions, and effects.
In principle, all preconditions for all scenes of all story modules
are evaluated continuously to determine the actions of the narrative
agents.  This is made computationally efficient using the Rete
algorithm.  Dialog is expressed using Curveship-style
templates~\cite{montfort:curveship}.

@@

Story modules can be inherited and extended, as shown in the figure.
This makes it easy to take an existing story and modify it to
better suit particular interests or a particular type of learner, or
just to add variety.  This can lead to conflicts: multiple versions of
a story, or multiple variant elements within a story, may have their
preconditions satisfied simultaneously.

We resolve conflicts...

@@

<category>
 <pattern>NORTH</pattern>
 <template>
  <javascript>FireEvent(Chat, "North")</javascript>
 </template>
</category>
</aiml>
\end{verbatim}
\caption{An AIML fragment which recognizes the commands ``N'',
  ``North'', and ``Go North''.}\label{fig:aiml}
\end{figure}
% show fragment
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.  Figure~\ref{fig:aiml}
illustrates how this can be used to implement a traditional
text adventure interaction for a particular story module.
The child can write instructions such as, ``go north,'' ``go
east,'' or ``open door'' for the agent.  These trigger events
handled by the story module to move the agent through the
adventure---for example, by causing the activity to draw a
new location, causing the agent to describe it, and shifting the
AIML topic to enable vocabulary suitable for the new location.

>>>>

\subsection{Rule-based story modules}
\begin{figure}\small
\begin{verbatim}
{ world: { // id gives canonical URL for this module
    id: "/nell.laptop.org/chapters/alphabet-book",
    scene: "opening", // initial scene
    inherit_from: "/nell.laptop.org/core/drawing_activity",
  }, scenes: [
    { id: "opening", subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Hello! Would (o/2) like to hear a story?",
      group: "/nell.laptop.org/core/story-intro-group",
    },
    { id: "story-1",
      cond: rule`At(opening) && Event(Chat, Yes)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Once upon a time..." },
    ...
    { id: "r-drawn",
      cond: rule`Event(Activity, DrawR)`,
      subject: "actors/princess",
      verb: "speak", object: "actors/user",
      text: "Very good, (o/3)!",
      action: js`startAnimation("r-runs")` },
    ...
  ], ... }
\end{verbatim}
\caption{Excerpt of story module.  JSON syntax has been extended with
quasiquote~\cite{quasiquote}.}\label{fig:rules}
\end{figure}
Narrative is the core of the Nell interface.  It is important that
Nell's story be modular and extensible.  We use a rule-based story
model with programmable conflict resolution to provide this ability.
Figure~\ref{fig:rules} shows an excerpt of the story module driving
the interaction described in the introduction to this paper.

Stories consist of a sequence of \textit{scenes}, in addition to
characters, locations, and activity information.  Scenes are ``goals''
in the STRIPS formalism~\cite{strips}, and consist of
action descriptions, preconditions, and effects.
In principle, all preconditions for all scenes of all story modules
are evaluated continuously to determine the actions of the narrative
agents.  This production rules system
is made computationally efficient using the Rete
algorithm.  Dialog is expressed using Curveship-style
templates~\cite{montfort:curveship}.

@@

Story modules can be inherited and extended, as shown in the figure.
This makes it easy to take an existing story and modify it to
better suit particular interests or a particular type of learner, or
just to add variety.  This can lead to conflicts: multiple versions of
a story, or variant scenes within a story, may have their
preconditions satisfied simultaneously.

Conflicts are resolved by consulting a \textit{group} associated with
each scene.  Each group has a default priority and names a scene to be
invoked when a conflict exists; the actions associated with that scene
will temporarily alter the priorities of the conflicting scenes to
resolve the conflict.  For example, the \textit{story-intro-group}
cited in the example defaults to a lower priority, such that
continuing a story in progress is preferred to beginning a new scene.
The resolution scene for \textit{story-intro-group} may be part of its
own story: for example, it may bring us to the controls of a magic airship
which allows us to pilot between locations corresponding to the
various story continuations.  Selecting a destination location bumps
the priority of the given scene allowing us to transition to it
without further conflict.

Other rule groups may invoke random selection, select between stories
based on child preferences or curriculum, or (during development) flag
an error to highlight an unexpected ambiguity.

@@

<category>
 <pattern>NORTH</pattern>
 <template>
  <javascript>FireEvent(Chat, "North")</javascript>
 </template>
</category>
</aiml>
\end{verbatim}
\caption{An AIML fragment which recognizes the commands ``N'',
  ``North'', and ``Go North''. The \verb|<srai>| tag invokes the
  recognizer recursively.}\label{fig:aiml}
\end{figure}
% show fragment
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.  Figure~\ref{fig:aiml}
illustrates how this can be used to implement a traditional
text adventure interaction for a particular story module.
The child can write instructions such as, ``go north,'' ``go
east,'' or ``open door'' for the agent.  These trigger events
handled by the story module to move the agent through the
adventure---for example, by causing the activity to draw a
new location, causing the agent to describe it, and shifting the
AIML topic to enable vocabulary suitable for the new location.
